```{r, echo=FALSE, warning=FALSE, message=FALSE, results='hide'}

source(paste0("../", code_dir, "params.R"))

```

Is international trade free and fair? For trade to be free, firms must not face government-imposed burdens to foreign market access. I refer to these burdens as policy barriers to trade. For trade to be fair, any policy barriers that do exist must treat products from all origin countries equally.^[Of course, there are many competing conceptions of what a free and fair international trading system should look like. These are the definitions of free and fair I use here.]

Examining tariff rates produces a qualified "yes," on both counts. Despite recent threats to the world trading system,^[See Bown, Chad P. ["Is the Global Trade System Broken?"](https://piie.com/commentary/op-eds/global-trade-system-broken) *Peterson Institute for International Economics*. 8 May 2018.] tariffs remain at historically low rates (less than five percent on most trade [@Baldwin2016]), suggesting trade is relatively free. Moreover, World Trade Organization (WTO) member countries, accounting for the vast majority of the world economy, commit to the principle of nondiscrimination (or most-favored-nation (MFN)) in tariff policy, applying the same tariff rates to the imports of all member countries. At first glance, adherence to this principle suggests international trade is also fair.

However, tariffs are but one instrument by which governments can influence the flow of trade. *Direct* barriers to trade are imposed at national borders or ports of entry. In addition to tariffs, governments also impose many non-tariff regulations on imports. Often referred to collectively as nontariff measures (NTMs), these regulations require that prospective importers comply with these price controls, quotas, quality and safety requirements, and other rules in order to access foreign markets.^[For studies of these kinds of barriers, see @Mansfield1995b; @Lee1997; @Gawande1999; @Kono2006; @Rickard2012; @Maggi2018.]

*Indirect*, or "behind-the-border", barriers are economic policies not assessed at the border that nevertheless disproportionately affect imported goods. Government procurement rules often explicitly privilege domestic suppliers, resulting in increased domestic purchases and reduced imports [@Evenett2004; @Kono2014]. Excise taxes, while implemented uniformly on a single good, may primarily fall on imports if targeted at goods with high foreign content.^[Sin taxes on alcohol and cigarettes might distort trade if these products are generally imported.] Subsidies and tax credits made available to domestic firms allow less productive firms to survive, supplanting importers in home markets and reducing trade. The burden of complying with health, safety, and environmental regulations may also fall disproportionately on foreign firms, reducing their sales and distorting trade.

All of these instruments can in principle be targeted to generate *de facto* discrimation. For example, the MFN principle is enforced at the tariff line level, allowing importers to target duties at products exported by specific countries, without running afoul of WTO rules. Through high agricultural duties, the United States, Europe, and Japan effectively discriminate against the developing world, which specializes in the production of these products [@Anderson2005]. NTMs and behind-the-border barriers can produce effective discrimination in the same manner.

Even armed with data on all such trade-distorting policy instruments, estimating the magnitude of aggregate policy barriers to trade would be challenging. Here, I propose and implement a new method to estimate policy barriers to trade with minimal data requirements. I construct a parsimonious model of international trade subject to frictions, following @Eaton2002.^[Theirs is a Ricardian model, in which the basis for trade emerges from differences in technologies across countries.] I show that the magnitude of trade frictions between two countries $i$ and $j$ is related by the theoretical model to price levels in both countries, trade flows between them, and the market shares of domestic producers in home markets. I then decompose these barriers into their economic (transportation costs) and political (policy barriers) components. Finally, I calibrate this relationship to the data on prices, trade, and freight costs in `r Y`.

The intuition underlying the model is straightforward. Cross-national price gaps inform about the existence of arbitrage opportunities, and imply that large trade flows should exist from countries with low prices toward those with high prices. The extent to which these flows are realized in the data informs about the magnitude of trade costs. If the cost of freight between countries is known, then the component of these costs than cannot be attributed to purely economic frictions can be independently identified. The remaining "missing trade" is attributed to the existence of policy distortions, broadly defined.

The logic behind the approach employed here is also articulated in @Leamer1988. If consumers are homogenous across countries, they will consume the same basket of goods when trade is frictionless (and prices equalize across markets).^[Empirical studies of trade rely heavily on the (dubious) assumption of consumer homogeneity. For a prominent counterexample, see @Fajgelbaum2016. I hold consumers' preferences over tradable goods constant, but allow for heterogeneity in consumers' taste for tradable versus nontradable goods.] Observed heterogeneity in consumption baskets is then informative about the magnitude of trade frictions. Leveraging advances in the structural gravity literature, I am able to empirically connect Leamer's basic insight more tightly to theory.

The results point to far more policy distortion and effective discrimination than would be inferred from the tariff data. Tariff equivalents of implied policy barriers are generically more than an order of magnitude larger than observed tariffs. Moreover, exporters in subset of favored countries enjoy far superior market access conditions than their peers in unfavored countries.

```{r triIncome, echo=FALSE, warning=FALSE, message=FALSE, fig.cap = "Tariff rates (left) and structural trade restrictiveness (right) against GDP per capita \\label{fig:triIncome}", fig.height=5, dpi=300}

source(paste0("../", figs_dir, "triIncome.R"))

library(patchwork)
tarIncome + triIncome

```

The trade policy openness attributed to developed countries also depends strongly on the metric used to evaluate openness.^[See @Rodriguez2000, @Dollar2004, and @Tavares2008 for discussions of this phenomenon.] As shown in Figure \ref{fig:triIncome}, there is a negative association between economic development (per capita GDP) and applied tariff rates. This relationship is reversed if trade policy restrictiveness is measured as proposed here. Countries with higher per capita incomes tend to have higher Trade Restrictiveness Indices.^[See Equation \ref{eq:tri}.] This is consistent with @Kono2006 and @Queralt2015, which suggest that developed countries offset tariff reductions with increases in non-tariff direct barriers and (potentially distortionary) domestic taxes.

This paper is most closely related to the international economics literature on the estimation of trade costs, beginning with @Anderson2004. The particular methodology adopted here draws on several studies that link price gaps to these trade costs [@Eaton2002; @Waugh2010; @Simonovska2014; @Sposi2015; @Waugh2016]. I build on these studies by disentangling policy barriers to trade and freight costs, and connecting the implied policy barriers to observable trade policy instruments. A parallel literature focuses on the estimation of trade costs under the assumption that they are symmetric [@Head2001; @Novy2013].^[@Bergstrand2013 provide an alternative method to estimate the asymmetric barriers targeted here.] While transportation costs may be nearly symmetric, policy barriers are less likely to be [@Kono2008; @Tavares2008]. Such estimates therefore average over meaningful policy heterogeneity.

The paper is also related to efforts to use observable barriers to trade to construct indices of trade openness [@Sachs1995; @Anderson1996; @Kee2009]. These observable barriers may be a non-random sample from the universe of protectionist instruments, however. Here, I take advantage of the structure of the theoretical model to infer the magnitude of policy barriers from the price and trade data, rather than attempting to quantify observable barriers. @Hiscox2002 construct country-level measures of aggregate trade openness using a fixed effects approach. @Martini2018 constructs industry-level measures of trade restrictiveness, under the assumption that policy barriers are nondiscriminatory within industry. I sacrifice industry-level granularity in order to assess discrimination in the international trade policy regime.

The fields of comparative and international political economy rely heavily on imperfect measures of trade protectionism. Political economic theories of protectionism generally relates primitives of the economic and political environment to a government's choice of trade policy, broadly construed. In evaluating these theories, however, researchers generally resort to examining observable barriers to trade, such as applied tariff rates, NTM coverage ratios, or simply the volume of trade.^[For a few examples, see @Goldberg1999; @Mansfield2000b; @Milner2005; @Tavares2008; @Kono2009; @Gawande2009; @Betz2017; @Barari2019.] The measure constructed here is arguably closer to the theoretical quantity of interest of many of these studies.

The broad policy barriers recovered here are also the objects that governments seek to influence in international negotiations, particularly in today's era in which tariffs rates are historically low.^[For example, Trans Pacific Partnership (TPP) negotiations focused overwhelmingly on non-tariff liberalization efforts. Fergusson, Ian F. and Brock R. Williams. ["The Trans-Pacific Partnership (TPP): Key Provisions and Issues for Congress."](https://fas.org/sgp/crs/row/R44489.pdf) 14 June, 2016. Congressional Research Service.] Governments desire foreign market access for the firms whose interests they represent [@Gawande2009; @Ossa2011; @Ossa2012]. Acquiring foreign market access requires dismantling policy barriers to trade, direct and indirect. This places governments in a complex multilateral bargaining game that has attracted the attention of many studies.^[See, for example, @Hirschman1945; @Pollins1989b; @Gowa1993; @Milner1997; @Aghion2007; @Head2010; @Antras2011; @Dube2011; @Berger2013; @Ossa2014.] Evaluating and assessing the outcomes of this game requires measurement of its outcomes -- governments' trade policy choices.

Finally, many argue that international institutions, the WTO and its predecessor General Agreements on Tariffs and Trade (GATT) in particular, structure this bargaining game in important ways [@Bagwell1999; @Maggi1999; @Steinberg2002; @Davis2006; @Carnegie2014; @Bagwell2018b]. GATT signatories committed in principle to convert protective policy measures into tariff-equivalents and subsequently negotiated primarily over tariff barriers [@Bagwell2004]. Theories of international trade institutions generally take this commitment seriously, assuming commitments to reduce tariffs cannot be subsequently "undone" through the implementation of non-tariff or behind-the-border barriers to trade. Statements about the efficacy of the principles of reciprocity and nondiscrimination in achieving efficient outcomes rest on this premise.

I proceed in three steps. The next section specifies a model of international trade and demonstrates how it relates observables to the magnitude of trade policy distortions. I then discuss the data that I use to calibrate the model. Finally, I present the results of this exercise and discuss their implications for the question posed at the beginning of this paper -- is international trade free and fair?
