## Data and Measurement {id="measurement"}

$$
\tau_{ij} = \left( \frac{\lambda_{ij}}{\lambda_{jj}} \right)^{-\frac{1}{\theta}} \frac{P_i}{P_j} \frac{1}{\delta_{ij}}
$$

## Data and Measurement

$$
\tau_{ij} = \left( \color{bcOrange} \frac{\lambda_{ij}}{\lambda_{jj}} \color{black} \right)^{-\frac{1}{\theta}} \frac{P_i}{P_j} \frac{1}{\delta_{ij}}
$$

- $\color{bcOrange} \frac{\lambda_{ij}}{\lambda_{jj}}$ - relative expenditure shares
  + Data: trade flows and national accounts

## Data and Measurement

$$
\tau_{ij} = \left( \frac{\lambda_{ij}}{\lambda_{jj}} \right)^{-\frac{1}{\theta}} \color{bcOrange} \frac{P_i}{P_j} \color{black} \frac{1}{\delta_{ij}}
$$

- $\frac{\lambda_{ij}}{\lambda_{jj}}$ - relative expenditure shares
  + Data: trade flows and national accounts
- $\color{bcOrange} \frac{P_i}{P_j}$ - relative price levels
  + Data: International Comparison Program (World Bank), good prices and exchange rates
  + Estimate preference weights, $\bm{\alpha}$

## Data and Measurement

$$
\tau_{ij} = \left( \frac{\lambda_{ij}}{\lambda_{jj}} \right)^{-\frac{1}{\theta}} \frac{P_i}{P_j} \color{bcOrange}  \frac{1}{\delta_{ij}}
$$

- $\frac{\lambda_{ij}}{\lambda_{jj}}$ - relative expenditure shares
  + Data: trade flows and national accounts
- $\frac{P_i}{P_j}$ - relative price levels
  + Data: International Comparison Program (World Bank), good prices and exchange rates
  + Estimate preference weights, $\bm{\alpha}$
- $\color{bcOrange} \delta_{ij}(\bm{Z}_{ij})$
  + $\bm{Z}_{ij}$ dyadic geographic covariates
  + Data (I): freight costs for shipments to United States, Australia, sample of maritime trade (OECD)
  + Data (II): transportation mode shares for United States, Australia, Japan, European Union, and Brazil
  + Estimate function $\delta_{ij}(\bm{Z}_{ij})$ to predict freight costs out of sample

## Data and Measurement

$$
\tau_{ij} = \left( \frac{\lambda_{ij}}{\lambda_{jj}} \right)^{-\frac{1}{\theta}} \frac{P_i}{P_j} \frac{1}{\delta_{ij}}
$$

- $\frac{\lambda_{ij}}{\lambda_{jj}}$ - relative expenditure shares
  + Data: trade flows and national accounts
- $\frac{P_i}{P_j}$ - relative price levels
  + Data: International Comparison Program (World Bank), good prices and exchange rates
  + Estimate preference weights, $\bm{\alpha}$
- $\delta_{ij}(\bm{Z}_{ij})$
  + $\bm{Z}_{ij}$ dyadic geographic covariates
  + Data (I): freight costs for shipments to United States, Australia, sample of maritime trade (OECD)
  + Data (II): transportation mode shares for United States, Australia, Japan, European Union, and Brazil
  + Estimate function $\delta_{ij}(\bm{Z}_{ij})$ to predict freight costs out of sample
- Year: 2011

. . . 

**Structural Paramters**

$$
\bm{\theta} = \left\{ \sigma, \theta, \bm{\nu} \right\}
$$

## Expenditure Shares $\frac{\hat{\lambda}_{ij}}{\hat{\lambda}_{jj}}$

- Free on board (f.o.b.): value of goods pre-shipment
- Cost, insurance, and freight (c.i.f.): value of goods at port of entry

$$
\lambda_{ij} = \tau_{ij} \lambda_{ij}^{\text{cif}} = \tau_{ij} \delta_{ij} \lambda_{ij}^{\text{fob}}
$$

$$
\lambda_{jj}(\bm{\tau}) = \left( 1 - \sum_{k \neq j} \tau_{jk} \lambda_{jk}^{\text{cif}} \right)
$$

## Expenditure Shares $\frac{\hat{\lambda}_{ij}}{\hat{\lambda}_{jj}}$

- Free on board (f.o.b.): value of goods pre-shipment
- Cost, insurance, and freight (c.i.f.): value of goods at port of entry

$$
\lambda_{ij} = \tau_{ij} \lambda_{ij}^{\text{cif}} = \tau_{ij} \delta_{ij} \lambda_{ij}^{\text{fob}}
$$

$$
\lambda_{jj}(\bm{\tau}) = \left( 1 - \sum_{k \neq j} \tau_{jk} \lambda_{jk}^{\text{cif}} \right)
$$

$$
\color{bcOrange} \lambda_{jk}^{\text{cif}} \color{black} = \frac{X_{jk}^{\text{cif}}}{E_j}
$$

- $X_{jk}^{\text{cif}} = \delta_{ij}(\bm{Z}_{ij}) X_{ij}^{\text{fob}}$
- $X_{ij}^{\text{fob}}$ - COMTRADE
- $E_j = \text{GC}_j - (1 - \nu_j) \text{GDP}_j$ - national accounts (OECD and WIOD)
- $\nu_j$ - share of income spent on tradables (ICP)

## Modified Estimating Equation

<hr style="height:8em; visibility:hidden;" />

$$
\tau_{ij} = \left( \frac{\lambda_{ij}^{\text{cif}}}{ \color{bcOrange} \lambda_{jj}(\bm{\tau}_j) \color{black} } \right)^{-\frac{1}{\theta + 1}} \left( \frac{\hat{P}_i}{\hat{P}_j} \right)^{\frac{\theta}{\theta+1}} \left( \frac{1}{\delta_{ij}(\bm{Z}_{ij})} \right)^{\frac{\theta}{\theta+1}}
$$

## Preliminary Evidence of Significant Barriers {id="shares"}

```{r shares1, echo=FALSE, warning=FALSE, message=FALSE, fig.height=5.5, dpi=300}

source(paste0("../", figs_dir, "shares.R"))

sharesCF + plot_spacer()

```

## Preliminary Evidence of Significant Barriers

```{r shares2, echo=FALSE, warning=FALSE, message=FALSE, fig.height=5.5, dpi=300}

sharesCF + sharesF

```

## Price Levels $\frac{\hat{P}_i}{\hat{P}_j}$

**Empirical Price Index**

$$
\hat{P}_i(\bm{w}) = \left( \int_{\omega \in [0,1]} \color{bcOrange} \hat{\alpha}_{h(\omega)} \color{black} p_i^\star(\omega)^{1 - \sigma} d \omega \right)^{\frac{1}{1 - \sigma}}
$$

**Good Categories**

$$
h : \Omega \rightarrow \mathcal{K}
$$

. . .

**Assumption:** $p_i(\omega) = p_i(\omega^\prime) = p_{ik}$ for all $\omega, \omega^\prime \in \Omega_k$

- $p_{ik}$ - price in country $i$ of good ICP basic heading $k$ (e.g. Rice)

## Price Levels $\frac{\hat{P}_i}{\hat{P}_j}$

**Demand**

$$
x_{ik} = \int_{\omega \in \Omega_k} \alpha_k p_{ik}^{1 - \sigma} E_i^q P_i^{\sigma - 1} d \omega
$$

. . .

**Normalization**

- $\alpha_{0} = 1$

$$
\frac{\lambda_{ik}}{\lambda_{i0}} = \alpha_k \left( \frac{p_{ik}}{p_{i0}} \right)^{1 - \sigma}
$$

- $\frac{\lambda_{ik}}{\lambda_{i0}}$ - basic heading expenditures (ICP)

. . .

**Estimation**

$$
\ln \Delta \lambda_{ik} = \ln \alpha_k + (1 - \sigma) \ln \Delta p_{ik} + \ln \epsilon_{ik}
$$

## Price Levels (Estimates) {id="priceLevels"}

```{r prices, echo=FALSE, warning=FALSE, message=FALSE, fig.height=5, dpi=300}

source(paste0("../", figs_dir, "P.R"))

plotP + plotTshare

```

## Freight Costs $\delta_{ij}(\bm{Z}_{ij})$

- Products $m \in \left\{ 1, ..., M \right\}$
- Transportation modes $o \in \left\{ 1, ..., O \right\}$
- $\zeta_{ij}^{m o}$ share of product $m$ transported by mode $o$
  + United States, Australia, Japan, European Union, and Brazil
- $\delta_{ij}^{m o}$
  + United States, Australia, sample of maritime shipments (OECD)

. . .

**Mode and Cost Functions**

$$
g: \left\{ \boldsymbol{Z}_{ij}, d^m \right\} \rightarrow \delta_{ij}^{m o}
$$
$$
h: \left\{ \boldsymbol{Z}_{ij}, d^m \right\} \rightarrow \zeta_{ij}^{m o}
$$

. . .

**Aggregate Freight Costs**

$$
\hat{\delta}_{ij} \left( \boldsymbol{Z}_{ij}, \boldsymbol{d}_{ij} \right) = \frac{1}{X_{ij}} \sum_{m = 1}^M x_{ij}^m \sum_{o=1}^O g \left( \boldsymbol{Z}_{ij}, d^m \right) h \left( \boldsymbol{Z}_{ij}, d^m \right)
$$

## Freight Costs (Cross Validation) {id="freightCosts"}

```{r freightCV, echo=FALSE, warning=FALSE, message=FALSE, fig.height=5, fig.cap="Out of sample (black): New Zealand and Chile", dpi=300}

source(paste0("../", figs_dir, "freight.R"))

freightCV + theme(aspect.ratio=1)

```

## Parameter Calibration

$$
\bm{\theta} = \left\{ \sigma, \theta, \bm{\nu} \right\}
$$

- Trade elasticity: $\sigma - 1 = \theta =$ `r theta`
  + @Simonovska2014, @Head2014

. . .

- $\bm{\nu}$ - tradable expenditure share (ICP)
