# Free and Fair? {id="introduction"}

<div class="fragment">

> "We believe that trade must be fair and reciprocal. The United States will not be taken advantage of any longer." <br/> $\qquad$ -- Donald Trump, Feb. 2019

</div>
<div class="fragment">

> "I’m going to have a trade prosecutor...to make sure every trade deal we’re already in is absolutely followed to the letter so nobody takes advantage of us." <br/> $\qquad$ -- Hillary Clinton, Oct. 2016

</div>
<div class="fragment">

> "Internationally, we are seeing a tendency toward protectionism and navel-gazing...What we need is trade that’s both fair and free." <br/> $\qquad$ -- Shinzo Abe, Oct. 2016

</div>
<div class="fragment">

> "An open world is only worth it if the competition that takes place there is fair." <br/> $\qquad$ -- Emmanuel Macron, Sep. 2017

</div>
<div class="fragment">

- **Free:** Foreign firms enjoy same market access conditions as domestic competitors \pause

</div>
<div class="fragment">

- **Fair:** Barriers that do exist affect all trade partners symmetrically

</div>

## The Tariff System (GATT/WTO)

- **Free?** Applied tariff rates are low, ~5% on average [@Baldwin2016]
- **Fair?** WTO members (vast majority of world economy) commit to principle of non-discrimination (Most Favored Nation)

```{r tcostsF, echo=FALSE, warning=FALSE, message=FALSE, fig.height=4, dpi=300}

source(paste0("../", figs_dir, "tcosts.R"))

tCostsF + plot_spacer()

```


## The Tariff System (GATT/WTO)

- **Free?** Applied tariff rates are low, ~5% on average [@Baldwin2016]
- **Fair?** WTO members (vast majority of world economy) commit to principle of non-discrimination (Most Favored Nation)

```{r tcostsFT, echo=FALSE, warning=FALSE, message=FALSE, fig.height=4, dpi=300}

tCostsF + tCostsFT

```

## Varieties of Barriers

**Direct**

- Tariffs
- Non-Tariff Measures (NTMs)
    + Price controls, quotas, health and safety regulations, technical barriers
    + @Mansfield1995b, @Lee1997, @Gawande1999, @Kono2006, @Rickard2012, @Maggi2018

. . .

**Indirect** (Behind-the-Border)

- Government procurement
    + @Evenett2004, @Kono2014
- Subsidies
- Excise taxes
- Regulations

. . .

**Effective Discrimination:** Target trade-disorting instruments to disproportionately affect disfavored trading partners.

- E.g. high agricultural duties by developed countries disproportionately harm developing countries

## Approach

**Trade Costs:** Policy distortions and transportation costs to access market $i$ in excess of those faced by firms in country $i$

. . .

**Overview**

1. Build model of international economy subject to trade costs [@Eaton2002]

. . .

2. Derive structural relationship linking trade costs to price levels, trade flows, market shares of home producers (observables)
    + @Anderson2003, @Waugh2010, @Simonovska2014, @Sposi2015, @Waugh2016

. . .

3. Decompose trade costs into economic (transportation costs) and political (policy barriers) components

. . .

4. Model transportation costs, combine with data on variables in (2) to separately estimate magnitude of policy barriers

. . .

$\tau_{ij} - 1 \geq 0$ -- ad valorem tariff equivalent of policy barriers gov $i$ imposes on gov $j$

## Estimand

<hr style="height:7em; visibility:hidden;" />

$$
\bm{\tau}_{N \times N} = \begin{pmatrix}
1 & \cdots & \tau_{1n} \\
\vdots & \ddots & \vdots \\
\tau_{n1} & \cdots & 1
\end{pmatrix}
$$

## Measurement Matters {id="triIncome"}

```{r triIncome1, echo=FALSE, warning=FALSE, message=FALSE, fig.height=5.5, dpi=300}

source(paste0("../", figs_dir, "triIncome.R"))

tarIncome + plot_spacer()

```

## Measurement Matters

```{r triIncome2, echo=FALSE, warning=FALSE, message=FALSE, fig.height=5.5, dpi=300}

tarIncome + triIncome

```