```{r, echo=FALSE, warning=FALSE, message=FALSE, results='hide'}

tarYval <- read_csv(paste0(otherdir, "tarYval.csv")) %>% pull(.)

trimaiY <- read_csv(paste0(resultsdir, "triMaiY.csv"))
gdp <- read_csv(paste0(cleandir, "gdp.csv"))
pop <- read_csv(paste0(cleandir, "pop.csv"))
tar <- read_csv(paste0(cleandir, "tarY.csv"))

correlates <- read_csv(paste0(resultsdir, "/correlates.csv"))

delta <- read_csv(paste0(cleandir, "delta.csv")) %>% filter(year==Y)
tau <- read_csv(paste0(resultsdir, "tauY.csv"))
tar <- read_csv(paste0(cleandir, "tarY.csv"))

shares <- read_csv(paste0(cleandir, "shares.csv"))

tauHMY <- read_csv(paste0(resultsdir, "tauHMY.csv"))
tauHMYEUD <- read_csv(paste0(resultsdirEU, "tauY.csv"))

P <- read_csv(paste0(cleandir, "priceIndex.csv"))

freight <- read_csv(paste0(cleandir, "freight.csv"))

```


# Results {id="results"}

<hr style="height:3em; visibility:hidden;" />

$$
\bm{\tau}_{N \times N} = \begin{pmatrix}
1 & \cdots & \tau_{1n} \\
\vdots & \ddots & \vdots \\
\tau_{n1} & \cdots & 1
\end{pmatrix}
$$

<hr style="height:2em; visibility:hidden;" />

$$
\tau_{ij} = \left( \frac{\lambda_{ij}^{\text{cif}}}{ \lambda_{jj}(\bm{\tau}_j) } \right)^{-\frac{1}{\theta + 1}} \left( \frac{\hat{P}_i}{\hat{P}_j} \right)^{\frac{\theta}{\theta+1}} \left( \frac{1}{\delta_{ij}(\bm{Z}_{ij})} \right)^{\frac{\theta}{\theta+1}}
$$

## Results


```{r hm1, echo=FALSE, warning=FALSE, message=FALSE, fig.height=4.5, dpi=300}

EUHM <- FALSE
cluster <- FALSE
highlight <- NULL

source(paste0("../", figs_dir, "hm.R"))
hm

```

## Results


```{r hmUSA, echo=FALSE, warning=FALSE, message=FALSE, fig.height=4.5, dpi=300}

EUHM <- FALSE
cluster <- FALSE
highlight <- "USA"

source(paste0("../", figs_dir, "hm.R"))
hm

```

. . .

- U.S. favored partners: Canada (`r tau %>% filter(j_iso3=="USA", i_iso3=="CAN") %>% pull(tau) %>% round(2) %>% -1 %>% "*"(100)`% ad valorem), Japan (`r tau %>% filter(j_iso3=="USA", i_iso3=="JPN") %>% pull(tau) %>% round(2) %>% -1 %>% "*"(100)`%), EU (`r tau %>% filter(j_iso3=="USA", i_iso3=="EU") %>% pull(tau) %>% round(2) %>% -1 %>% "*"(100)`%)
- U.S. disfavored partners: Peru (`r tau %>% filter(j_iso3=="USA", i_iso3=="PER") %>% pull(tau) %>% round(2) %>% -1 %>% "*"(100)`%) and Vietnam (`r tau %>% filter(j_iso3=="USA", i_iso3=="VNM") %>% pull(tau) %>% round(2) %>% - 1 %>% "*"(100)`%)

## Results (Free?)

```{r tCostsPlot1, echo=FALSE, warning=FALSE, message=FALSE, fig.height=5.5, dpi=300}

source(paste0("../", figs_dir, "tcosts.R"))

tCostsFT + theme(aspect.ratio=1)

```

## Results (Free?)

```{r tCostsPlot2, echo=FALSE, warning=FALSE, message=FALSE, fig.height=5.5, dpi=300}

tCostsFTPbase + theme(aspect.ratio=1)

```

## Results (Free?) {id="magnitudes"}

```{r tCostsPlot3, echo=FALSE, warning=FALSE, message=FALSE, fig.height=5.5, dpi=300}

tCostsFTP + theme(aspect.ratio=1)

```

## Trade Restrictiveness and Market Access 

**Trade Restrictiveness Index**

$$
\text{TRI}_i = \frac{1}{\sum_{j \neq i} E_j} \sum_{j \neq i} \tau_{ij} E_j
$$

**Market Access Index**

$$
\text{MAI}_j = \frac{1}{\sum_{i \neq j} E_i} \sum_{i \neq j} \tau_{ij} E_i
$$

## Results (Fair?) {id="TRI"}

```{r triincome, echo=FALSE, warning=FALSE, message=FALSE, fig.height=5, dpi=300}

source(paste0("../", figs_dir, "triIncome.R"))

triIncome + theme(aspect.ratio=1)

```

## Results (Fair?) {id="MAI"}

```{r maigdppc, echo=FALSE, warning=FALSE, message=FALSE, fig.height=5, dpi=300}

source(paste0("../", figs_dir, "maigdppc.R"))

maigdppcPlot + theme(aspect.ratio=1)

```